Pool Model - A unified Mathematical\\ Modelling Framework for Food Science

### 2025-10-30 paper/pool_model.tex
adjust to journal latex template

\documentclass[10pt,A4paper]{article}
\usepackage[utf8]{inputenc}

\setlength{\tabcolsep}{1.3pt}
\setlength{\columnsep}{0.4cm}
\usepackage{acronym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[
    backend=biber,
    sorting=none,
    style=numeric-comp,
    hyperref=true,
    maxnames=2,
    url=false,
    isbn=false,
    eprint=false,
    date=year,
    doi=false,
    giveninits=true
]{biblatex}
\usepackage{array}
\usepackage{bm}
\usepackage{booktabs}
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off, font=small]{caption}
\usepackage{diagbox}
\usepackage{float}
\usepackage[T1]{fontenc}
\usepackage[a4paper,inner=2cm,outer=2cm,top=2.5cm,bottom=2.5cm]{geometry}
\usepackage{graphicx}% Include figure files
\usepackage{here}
\usepackage{hyperref}
\usepackage{lipsum}
\usepackage{listings}
\usepackage{listings-rust}
\usepackage{mathtools}
\usepackage{mathrsfs}
\usepackage{multirow}
\usepackage{newfloat}
\usepackage{onimage}
\usepackage{siunitx}
\usepackage{subcaption}
\usepackage{tabularx}
\usepackage{tikz}
\usepackage[dvipsnames]{xcolor}
\usepackage{xpatch}

@@

% Add bibliography
\addbibresource{references.bib}

@@

\title{Pool Model - A Mathematical Framework\\ for Modelling Bacterial Consortia in Food Science}
\author{
    Polina Gaindrik\textsuperscript{1},
    Jonas Pleyer\textsuperscript{1}
    and Christian Fleck\textsuperscript{1}
}

\date{
    \footnotesize
    \textsuperscript{\textbf{1}}Freiburg Center for Data Analysis and Modeling and AI
}

\begin{document}
\maketitle

\begin{abstract}
    Bacterial consortia are complex systems which are often comprised of multiple distinct species.
    Their dynamics on a population-based level often result from interactions that lead to
    inter-species cooperation or competition.
    Various empirical and inactivation models have been conceived and applied to problems related to
    food microbiology.
    However, the utilized models live on the population level, omitting any spatial effects such as
    heterogeneity or diffusion.
    In this work we present a novel generalized framework, called "pool model" which uses multiple
    distinct pools that can hold bacterial states of distinct species or extracellular nutrient
    abundance.
    It allows us to describe these systems mechanistically as variations of an underlying, more
    fundamental model.
    To illustrate the usefulness of the model, multiple different examples are reported.
    Finally, we construct an \acl{abm} for a given pool model, showcasing that the resulting model
    arises naturally from the individual behaviour of many cellular agents.
    We further apply this \ac{abm} to study the role of spatial effects on the dynamcis of bacterial
    growth.
    % \textbf{Notes from Christian}
    % Different empirical and inactivation models applied in the estimation
    % of parameters and modelling of phenomena in food microbiology, gen-
    % eralized Lotka Voltera approach used to describe the interaction of mi-
    % croorganisms, mechanistic models utilized to provide more insight into
    % the underlying dynamic can all be unified as variations of a new model
    % known as three pool model. Compared against the aforementioned
    % methods which are applicable in specific applications with some limi-
    % tations, the present model turns out to be less complex with general
    % applicability, low number of parameters and preferably to explain the
    % mechanism of the phenomenon of interest. To illustrate the generality
    % of the model, three different examples are reported.
\end{abstract}

\textbf{Keywords:}
Bacterial Consortia, Competition and Cooperation, Mechanistic Modeling, Spatial Effects

\newpage
\tableofcontents

\twocolumn

\newpage
%
%
%
%
% ================================================================================================
\section{Introduction}

@@

@@

\printbibliography

>>>>

% \documentclass[10pt,A4paper]{article}
\documentclass[10pt,twocolumn,5p]{elsarticle}
\usepackage[utf8]{inputenc}

\setlength{\tabcolsep}{1.3pt}
\setlength{\columnsep}{0.4cm}
\usepackage{acronym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{array}
\usepackage{bm}
\usepackage{booktabs}
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off, font=small]{caption}
\usepackage{diagbox}
\usepackage{float}
\usepackage[T1]{fontenc}
% \usepackage[a4paper,inner=2cm,outer=2cm,top=2.5cm,bottom=2.5cm]{geometry}
\usepackage{graphicx}% Include figure files
\usepackage{here}
\usepackage{hyperref}
\usepackage{lipsum}
\usepackage{listings}
\usepackage{listings-rust}
\usepackage{mathtools}
\usepackage{mathrsfs}
\usepackage{multirow}
\usepackage{newfloat}
\usepackage{onimage}
\usepackage{siunitx}
\usepackage{subcaption}
\usepackage{tabularx}
\usepackage{tikz}
\usepackage[dvipsnames]{xcolor}
\usepackage{xpatch}

@@

\bibliographystyle{abbrvnat}

@@

\tnotetext[t1]{TODO}
\tnotetext[t2]{TODO}

\title{Pool Model - A unified Mathematical\\ Modelling Framework for Food Science}
\author[1]{Polina Gaindrik}
\author[1]{Jonas Pleyer\fnref{fn1}}
\author[1]{Christian Fleck\corref{cor1}}
\ead{christian.fleck@fdm.uni-freiburg.de}

\cortext[cor1]{Corresponding Author}
\fntext[fn1]{jonas.pleyer.org}
\affiliation[1]{
    organization={Freiburg Center for Data Analysis and Modeling},
    addressline={Ernst-Zermelo Str. 1},
    city={Freiburg im Breisgau},
    citysep={},
    postcode={79104},
    country={Germany},
}

\date{
    \footnotesize
    \textsuperscript{\textbf{1}}
}

\begin{document}

\begin{abstract}
    Bacterial consortia are complex systems which are often comprised of multiple distinct species.
    Their dynamics on a population-based level often result from interactions that lead to
    inter-species cooperation or competition.
    Various empirical and inactivation models have been conceived and applied to problems related to
    food microbiology.
    However, the utilized models live on the population level, omitting any spatial effects such as
    heterogeneity or diffusion.
    In this work we present a novel generalized framework, called "pool model" which uses multiple
    distinct pools that can hold bacterial states of distinct species or extracellular nutrient
    abundance.
    It allows us to describe these systems mechanistically as variations of an underlying, more
    fundamental model.
    To illustrate the usefulness of the model, multiple different examples are reported.
    Finally, we construct an \acl{abm} for a given pool model, showcasing that the resulting model
    arises naturally from the individual behaviour of many cellular agents.
    We further apply this \ac{abm} to study the role of spatial effects on the dynamcis of bacterial
    growth.
\end{abstract}

\maketitle

\begin{graphicalabstract}
% TODO
\end{graphicalabstract}

\begin{keyword}
Bacterial Consortia \sep Competition and Cooperation \sep Mechanistic Modeling \sep Spatial Effects
\end{keyword}
%
% ================================================================================================
\section{Introduction}

@@

% TODO change the name "death pool"?

@@

% \printbibliography
\bibliography{references.bib}
